\documentclass[]{article}
% EJ latex definitions and style
\def\tightlist{}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\red{\color{red}}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\date{}

\begin{document}

\subsection{Week 2 Notes (5/31/18)}\label{week-2-notes-53118}

During our meeting:

\begin{itemize}
\tightlist
\item
  Any questions about chapter 1 of Wiggins?
\item
  Review worksheet and answers
\item
  Review/walkthrough code and figures
\item
  Discuss nondimensionalization of gLV equations
\item
  Discuss Lyapunov functions for dynamical systems
\end{itemize}

For next week:

\begin{itemize}
\tightlist
\item
  Read chapter 2 of Wiggins
\item
  Show that 2D gLV equations can be nondimensionalized so that

  \begin{align} \begin{split}
      \frac{\text{d}x_a}{\text{d}t} &= x_a (\mu_a - x_a + M_{ab} x_b) \\
      \frac{\text{d}x_b}{\text{d}t} &= x_b (\mu_b + M_{ba}x_a - x_b),
      \label{eq1}
  \end{split} 
  \end{align}

  and that you may further nondimensionalize time by setting
  \(\mu_a \to 1\).
\item
  Show that the Lyapunov function given by Tang, Yuan and Ma in Phys.
  Rev.~E 2013 satisfies the Lyapunov conditions given in chapter 2 of
  Wiggins. In our notation, the Lyapunov function they provide is

  \begin{equation}
  \begin{split}
    V(x_a, x_b) &= M_{ba} x_a^2/2 + M_{ab} x_b^2/2 \\ 
    &\quad - M_{ba} \mu_a x_a - M_{ab} \mu_b x_b + M_{ab} M_{ba} x_a x_b
  \end{split}
  \end{equation}

  Some hints: 1) remember that \(\hat{x}_a\) and \(\hat{x}_b\) are
  ``directions'' (like \(\hat{x}\) and \(\hat{y}\)) in a 2-dimensional
  space, 2) remember that \(\dot{V} = \nabla V \cdot \dot{\textbf{x}}\),
  where \(\dot{\textbf{x}}\) is the vector form of the dynamical system,
  and 3) assume \(M_{ab} > 0\) and \(M_{ba} >  0\) (this condition must
  be satisfied in order for there to be two stable steady states). Note
  that this equation satifies the Lyapunov conditions except for the
  fact that \(V(\bar{x}) = 0\) for two different \(\bar{x}\),
  corresponding to the two stable steady states. For this reason, this
  is called a \textit{split Lyapunov function}.
\item
  Generalize your code for solving the 2-dimensional gLV equations so
  that it can uses numpy arrays (\texttt{np.array}). Hint: consider the
  commands \texttt{np.dot(np.diag(mu), Y)} and
  \texttt{np.dot(np.diag(np.dot(M, Y)), Y)} do, and compare them to the
  2-dimensional gLV equations.
\item
  Modify the interaction matrix so that it is time dependent. Use Eq.
  (1) above, and set \(\mu_a = \mu_b = 1\). Make your interaction matrix
  \(M\) time dependent, with \(M_{aa} = M_{bb} = -1\) for all time.
  Then, choose \(M_{ab}\) and \(M_{ba}\) so that

  \begin{equation}
    \begin{cases}
  M_{ab} = M_{ba} = -.5, & \text{if } t < 5 \text{ or } t > 20 \\
  M_{ab} = M_{ba} = -1.5, & \text{if } 5 <= t <= 20.
    \end{cases}
    \end{equation}

  To do this, make a function \texttt{def M(t): some stuff; return
    np.array([[0, 1 + t], [2*t, 4]])}. Then simulate this system
  starting from the initial condition (.1, .9) and ensure it does what
  you would expect.
\end{itemize}

\end{document}
